Diagnosising the sensitivity of grounding line flux to changes in sub-ice shelf melting

### 2019-04-08 draft_v1.tex
Initial Overleaf Import

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/mismip_larsenc.pdf}
\caption{Plan view of surface ice speeds for MISMIP+ (a) and Larsen C ice shelf (b). The white curves show the grounding lines.}
\label{mismip_larsenc}
\end{figure}

@@

Our experiments are conducted in a manner similar to those of \cite{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, the uniform, 2 km mesh implies that grid cell centers naturally align with these boxes. For the Larsen C ice shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we use both 2$\times$2 km and 10$\times$10 km square boxes in order to additionally investigate the impacts of spatially averaging perturbations over larger areas. For Larsen C, we only use 20$\times$20 km square boxes (i.e., as in \citet{reese2018}). Lastly, for the MISMIP+ 2 km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is likely lateraly buttressed) and $y>40$ km (one half of the ice shelf due to the symmetry about the centerline).

@@

Despite many different factors (ice flow directions, horizontal gradients of ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc) existing, we here mainly present the results of the stress fields and the distances between perturbations and GL, as they appears to correlate more closely to the sensitivity of GL flux change to ice shelf perturbations. Similar to  \cite{furst2016} (Eqn \ref{butN}), we calculate buttressing numbers ($N_b$) for 8 different stress components, i.e., normal stresses along ($\sigma_f$) and across flow ($\sigma_{af}$), first ($\sigma_{p1}$) and second ($\sigma_{p2}$) principle stresses, normal stress along the $x$ ($\sigma_{x}$) and $y$ direction ($\sigma_{y}$), shear stress ($\sigma_s$) and hoop stress ($\sigma_h$), and get their corresponding GL flux response numbers,
\begin{equation}
N_b=1-\frac{\sigma_{obj}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=0.5\left(1-{\rho_i}/{\rho_w}\right)gH$). $\rho_i$ (910 kg m$^3$) and $\rho_w$ (1028 kg m$^3$) are the density of ice and ocean water, respectively. Table 1 lists these 8 different buttressing number ($\sigma_{obj}$) definitions, where $n_f$, $n_{af}$, $n_x$ and $n_y$ are the unit vector components of the ice flow and coordinate direction, $H$ is the ice thickness, $u$ is the radial flow velocity and $r$ is the radial axis coordinate. Hoop stress arises from regions where ice flow is divergent, and is another source of ice shelf resistence except for compression and lateral shearing of ice flow \citep{pegler2012, wearing2016}.

\begin{table*}
\centering
\caption{Stress components used in the buttressing (Eqn \ref{butN})}
\label{stressDef}
\begin{tabular}{|c|c|c|c|}
\hline
$\sigma_{obj}$ & def. & $\sigma_{obj}$ & def. \\
\hline
$\sigma_f$ & ${\bf{n}}^T_f\cdot{\bf{\sigma}}\cdot{\bf{n}}_f$ & $\sigma_{af}$ & ${\bf{n}}^T_{af}\cdot{\bf{\sigma}}\cdot{\bf{n}}_{af}$ \\ 
\hline 
$\sigma_{x}$ & ${\bf{n}}^T_x\cdot{\bf{\sigma}}\cdot{\bf{n}}_x$ & $\sigma_{y}$ & ${\bf{n}}^T_y\cdot{\bf{\sigma}}\cdot{\bf{n}}_y$ \\ 
\hline
$\sigma_{p1}$ & $\frac{\sigma_{x}+\sigma_{y}}{2}+\sqrt{\left(\frac{\sigma_{x}-\sigma_{y}}{2}\right)^2+\sigma_{xy}^2}$ & $\sigma_{p2}$ & $\frac{\sigma_{x}+\sigma_{y}}{2}-\sqrt{\left(\frac{\sigma_{x}-\sigma_{y}}{2}\right)^2+\sigma_{xy}^2}$ \\ 
\hline 
$\sigma_s$ & ${\bf{n}}^T_f\cdot{\bf{\sigma}}\cdot{\bf{n}}_{af}$ & $\sigma_h$ & $\eta H\frac{\partial}{\partial r}\left(\frac{u}{r}\right)$ \\ 
\hline

\end{tabular}
\end{table*}

\section{Results}
\subsection{MISMIP+}
Before and after the perturbation, the GL velocity change clearly shows a spatial pattern (Fig \ref{stndVarFluxP}a). By calculating the standard deviation of the velocity change along GL for each 2 km perturbation experiment, we can see that the GL velocity change data is more dispersed for perturbation near GL, while for pertuirbations at the central region of the ice shelf, the corresponding GL velocity change is less variant along GL. If we look at Figure \ref{stndVarFluxP}b, the integrated GL flux is more contributed by the GL cells that close to the perturbation spot, i.e., the neighboring GL cells appears to have more flux contribution weight than those far away from the perperturbation location, and the perturbations near the ice shelf center tend to impact more uniformly on the GL dynamics.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/stndVarFluxP.pdf}
\caption{(a) The distribution of the standard variance of the ice velocity changes on the grounding line of MISMIP+ for all perturbation experiments (note we only do the perturbation experiments for half space in $y$ of MISMIP+); (b) We find five cells on the grounding line that are the nearset to the perturbation spot. The x-axis is their mean distances to the perturbation box. The y-axis shows their combined flux contribution in the total integrated grounding line flux.}
\label{stndVarFluxP}
\end{figure}

This can be further examined by looking at the relationship between GL flux response number ($N_r$) and the buttressing number ($N_b$) for the first principle stress ($\sigma_{p1}$) at different $dG$ and $dC$ (Fig \ref{diffdCdG}). $dG$ ($dC$) is the minimum distance between perturbation boxes and GL (calving front). When both $dG$ and $dC$ increase, perturbation boxes are farther away from GL and calving front, the velocity changes at GL become less dispersible and their relationship gets less ``noisy'' and tends to show linear trends. The possible explanaions can be found in the following ``Discussions'' section.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/diffdCdG.pdf}
\caption{The relationships between the groundling line flux response number ($N_r$) and the buttressing number ($N_b$) at the perturbation spots by different geometric filters of $dG$ and $dC$. Larger $dG$ and $dC$ give perturbation spots that are more centralized in the ice shelf. Smaller $dG$ and $dC$ mean the perturbation locations are either close to the grounding line or the calving front. The color bars show the values of $dG$ for each perturbation spot.}

\label{diffdCdG}
\end{figure}

In fact, $\sigma_{p1}$ appears to be the best metric for predicting such linear relationship, comparing to the other 7 stress components (Fig \ref{diffStressComp}). The $N_b$ for $\sigma_f$, $\sigma_{af}$, $\sigma_x$ and $\sigma_y$ show some weak correlations with GL flux response number $N_b$. The shear stress and hoop stress, however, have no clear evidences of relating to $N_r$. Note that for all plots in Figure \ref{diffStressComp}, we have removed the ``noisy'' perturbation spots using $dG$ and $dC$. It is suprising that $\sigma_{p1}$ shows better performance than $\sigma_{p2}$ and $\sigma_s$, as $\sigma_{p2}$ gives the maximum buttressing number and predicts well the passive ice regions in \cite{furst2016} and the shear stress supply the main large resistence force for divergent ice flow regions (most of the MISMIP+ ice shelf) \citep{wearing2016}.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/diffStressComp.pdf}
\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the MISMIP+ 2 km $\times$ 2 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
\label{diffStressComp}
\end{figure}

The buttressing number for $\sigma_{p1}$ also performs well for the 10 km $\times$ 10 km square box perturbation experiments (Fig \ref{diffStressCompAvg}), for which we average the stresses across all cells in each square box. In fact, except the shear stress $\sigma_s$, all correlations are improved compared to the 2 km $\times$ 2 km perturbation experiments. In particular, the hoop stress shows some improved predictability as well. But again, $\sigma_{p2}$ is still not preferable than $\sigma_{p1}$.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/diffStressCompAvg.pdf}
\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the MISMIP+ 10 km $\times$ 10 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
\label{diffStressCompAvg}
\end{figure}

\subsection{Realistic ice shelves (Larsen C)}

To validate if there are the same $N_b$-$N_r$ patterns as in the MISMIP+ case for realistic ice shelves, we apply the same analysis processes as above for the Larsen C ice shelf. Because the mesh resolution varies from GL to calving front, we apply 20 km $\times$ 20 km square boxes to the perturbation experiments and adjust the box areas by counting actual cell numbers that each box includes. In addition, due to the complex geometry (GL shape and ice rises) of Larsen C ice shelf, we adopt $dG=50$ km and $dC = 50$ km for better removing those near-GL noises.

From Figure \ref{diffStressCompAvgL}, we can also see good linear $N_b$-$N_r$ relations for most of the stress components, especially for $\sigma_{p1}$, similar to the MISMIP+ case. Again, the buttressing number for $\sigma_{p2}$ doesn't show better GL flux response predictivity than $\sigma_{p1}$, $\sigma_{f}$, $\sigma_{af}$, $\sigma_{x}$ and $\sigma_{y}$. Unlike MISMIP+, the hoop stress of Larsen C ice shelf has no clear evidence relating the GL flux response number.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/diffStressCompAvgL.pdf}
\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the Larsen C 20 km $\times$ 20 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
\label{diffStressCompAvgL}
\end{figure}

@@

\section{Discussions}

Based on the above discussion we hypothesize that the two main factors controlling how a local perturbation in ice shelf thickness leads to changes in ice flux across the GL are 1) the distance between a perturbation location and the GL and 2) the first principle stress at the perturbation location. Figure \ref{distPerturbComp} shows an example of the impact of the perturbation location on the velocity changes across the ice shelf and GL. Despite the complex patterns of velocity change around the perturbation spot, we can see two clear features: 1) the magnitude of velocity change decreases with distances away from the perturbation spot and 2) the velocity change pattern becomes different and more disordering when the perturbation  approaches to GL.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{figs/perturb_theta_velchange.pdf}
	\caption{An example of ice speed changes (\%) across the MISMIP+ ice shelf and grounding line after apply the perturbations at different locations (blue circles).}
	\label{distPerturbComp}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig1.pdf}
\caption{Plan view of steady-state surface ice speeds for MISMIP+ (a) and present-day surface ice speed for Larsen C ice shelf (b). The white curves show the grounding lines.}
\label{fig1}
\end{figure}

@@

Our experiments are conducted in a manner similar to those of \cite{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, the uniform, 2 km mesh implies that grid cell centers naturally align with these boxes. For the Larsen C ice shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we use 2$\times$2 km square boxes that are also the real cell size. For Larsen C, we only use 20$\times$20 km square boxes (i.e., as in \citet{reese2018}). Lastly, for the MISMIP+ 2 km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is likely lateraly buttressed) and $y>40$ km (one half of the ice shelf due to the symmetry about the centerline).

@@

Despite the existence of many different factors (ice flow directions, horizontal gradients of ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc), we here mainly present the results of the stress fields and the distances between perturbations and GL, as they appears to correlate more closely to the sensitivity of GL flux change to ice shelf perturbations. Similar to  \cite{furst2016} (Eqn \ref{butN}), we calculate buttressing numbers ($N_b$) as follows,
\begin{equation}
N_b=1-\frac{\sigma_{nn}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=0.5\left(1-{\rho_i}/{\rho_w}\right)gH$). $\rho_i$ (910 kg m$^3$) and $\rho_w$ (1028 kg m$^3$) are the density of ice and ocean water, respectively. $\sigma_{nn}$ is the normal stress along certain horizontal direction.

%\begin{table*}
%\centering
%\caption{Stress components used in the buttressing (Eqn \ref{butN})}
%\label{stressDef}
%\begin{tabular}{|c|c|c|c|}
%\hline
%$\sigma_{obj}$ & def. & $\sigma_{obj}$ & def. \\
%\hline
%$\sigma_f$ & ${\bf{n}}^T_f\cdot{\bf{\sigma}}\cdot{\bf{n}}_f$ & $\sigma_{af}$ & ${\bf{n}}^T_{af}\cdot{\bf{\sigma}}\cdot{\bf{n}}_{af}$ \\ 
%\hline 
%$\sigma_{x}$ & ${\bf{n}}^T_x\cdot{\bf{\sigma}}\cdot{\bf{n}}_x$ & $\sigma_{y}$ & ${\bf{n}}^T_y\cdot{\bf{\sigma}}\cdot{\bf{n}}_y$ \\ 
%\hline
%$\sigma_{p1}$ & $\frac{\sigma_{x}+\sigma_{y}}{2}+\sqrt{\left(\frac{\sigma_{x}-\sigma_{y}}{2}\right)^2+\sigma_{xy}^2}$ & $\sigma_{p2}$ & $\frac{\sigma_{x}+\sigma_{y}}{2}-\sqrt{\left(\frac{\sigma_{x}-\sigma_{y}}{2}\right)^2+\sigma_{xy}^2}$ \\ 
%\hline 
%$\sigma_s$ & ${\bf{n}}^T_f\cdot{\bf{\sigma}}\cdot{\bf{n}}_{af}$ & $\sigma_h$ & $\eta H\frac{\partial}{\partial r}\left(\frac{u}{r}\right)$ \\ 
%\hline
%
%\end{tabular}
%\end{table*}

\section{Results and Discussions}
\subsection{Linear relationship between buttressing ($N_b$) and GL flux responses ($N_r$)}
The decrease of ice shelf buttressing tends to induce the increase of GLF \citep{gagliardini2010}. Therefore, the highly buttressed regions are in general more sensitive to sub-ice shelf melting for  marine ice sheet dynamics. It is thus useful to understand more quantitatively the relationship between the buttressing ``strength'' (number) and the GLF  changes. Can we predict the changes of GLF simply by the buttressing number ($N_b$)? By calculating the buttressing number ($N_b$) along the first principle stress ($\sigma_{p1}$) direction, we can see very weak $N_b$-$N_r$ correlations from the results of all perturbation experiments (Fig \ref{fig2}b). However, if we do not consider the perturbation points that are weakly buttressed ($x>480$ km; near the open shelf region) and are close to GL (the minimum distances between GL and perturbation points are greater than 5 km), we can clearly see a strong linear $N_b$-$N_r$ regression (Fig \ref{fig2}c). A similar strong linearity can also be found if we apply a different filtering approach using the ice thickness gradient field (Fig \ref{fig2}a), i.e., we keep the perturbation points that have thickness gradients smaller than 0.0007 (Fig \ref{fig2}d). This nonlinearity feature is likely caused by the intensive shearing and complex ice flow mechanics near GL (a strong transition zone from relatively slow grounded ice flow to fast extensively floating ice flow). We can again see this non-linearity region in our adjoint sensitivity experiments (Fig. \ref{fig9} in section ``Adjoint sensitivity'').

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig2.pdf}
    \caption{(a) The magnitude of ice thickness gradient around GL (white curve) for the steady-state MISMIP+ geometry; (b) The relationship of $N_b$-$N_r$ for all perturbation points; (c) The relationship of $N_b$-$N_r$ for perturbation points that are $>$10 km away from GL; (d) The relationship of $N_b$-$N_r$ for perturbation points with thickness gradient $<$7$times$10$^{-3}$. The red (blue) dots in the insets are GL (perturbation) cells.}
\label{fig2}
\end{figure}

\subsection{The buttressing directions}
According to Equation \ref{butN}, the buttressing number $N_b$ is computed by the normal stress ($\sigma_{nn}$) along specific directions. Therefore, the buttressing at certain perturbation points can be various for different directions we apply. In \cite{furst2016} the buttressing along two significant directions (flow and second principle stress ($\sigma_{p2}$)) was analyzed, and found that the $\sigma_{p2}$ buttressing, which points to the most compression direction, has the maximum impacts on the ``passive'' ice shelf regions. Here we get the correlation numbers ($r^2$) between $N_r$ and $N_b$ for different directions with respect to the first principle stress and flow direction, respectively (Fig \ref{fig3}). Differently, we find that $N_b$ along the $\sigma_{p1}$ direction shows the best regression performance, whereas the $\sigma_{p2}$ appears to show the weakest correlations to $N_r$. This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$). From Figure \ref{fig3}b, we see that for around 50\% of the perturbation points, their flow directions are around 30--50 degree more than the $\sigma_{p1}$ directions, which is consistent with the phase differences in Figure \ref{fig3}a. If we add around 40 degree on top of the flow direction (blue curve), the new direction will be likely aligned closely with the $\sigma_{p2}$ direction, which is consequently pointing to the smallest $r^2$ number for the $\sigma_{p1}$ (red) curve. This indicates that the local maximum buttressing relating to $\sigma_{p2}$ is unnecessarily corresponding to the integrated instantaneous GLF responses.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig3.pdf}
    \caption{(a) The $N_b$-$N_r$ regression coefficients for each direction rotated anti-clock-wisely from the $\sigma_{p1}$ (red) and flow direction (blue); (b) The histogram of the angle differences between flow and $\sigma_{p1}$ directions. The perturbation points we apply here are the same as in Figure \ref{fig2}d.}
    \label{fig3}
\end{figure}

\subsection{Possible controlling factors}
Further clues of the impacts from different chosen directions can be found in Figure \ref{fig4}. We analyzed the number of maximum (light red) and minimum (blue) velocity changes around each perturbation point for the cases using flow (Fig \ref{fig4}a) and $\sigma_{p1}$ directions (Fig \ref{fig4}b). Figure \ref{fig4}a and c contain all perturbation points while Figures \ref{fig4}b and d only include the filtered perturbation points as in Figure \ref{fig2}d. For the flow direction results, most of the maximum (minimum) velocity change events occur along the flow (120 degree) direction. The $\sigma_{p1}$ direction results are more spread than of the flow direction results. However, it is still clear that most of the maximum (minimum) velocity change events are aligned near the first (second) principle stress direction, a supporting evidence for our previous findings.

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig4.pdf}
    \caption{Histograms of the frequency of the neighboring maximum (red) and minimum (blue) velocity change around all perturbation points (a and c) and around the same selected perturbation points (b and d) as Figure \ref{fig2}d ({\bf{TZ: this figure still needs double check!}})}
	\label{fig4}
\end{figure}

Another evidence is the standard deviation of the differences between the relative changes of stress and velocity ($\Psi$) along GL (Fig \ref{fig5}). 
\begin{equation}
\Psi = \frac{\sigma_{p}-\sigma_{c}}{\sigma_{c}} - \frac{u_{p}-u_{c}}{u_{c}},
\label{psi}
\end{equation}
where the subscripts $p$ and $c$ denote perturbation experiments and ctrl runs. $\sigma$ and $u$ denote the stress component and ice velocity, respectively. This is a measure of the consistency of velocity and stress variances along GL under certain perturbation experiments. Here we choose $\sigma_{p1}$ and $\sigma_{p2}$ for instances, i.e., we set $\sigma$ in Equation \ref{psi} to $\sigma_{p1}$ and $\sigma_{p2}$. Overall, the $\sigma_{p1}$ case (Fig \ref{fig5}a) shows smaller standard deviations, especially for perturbation points close to GL, than the $\sigma_{p2}$ case (Fig \ref{fig5}b). This possibly indicates that GLF are more relevant to the changes of $\sigma_{p1}$ along GL, compared to the $\sigma_{p2}$ case, which can be clearly found if we look at some specific perturbation example as shown in Figure \ref{fig6}.

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig5.pdf}
    \caption{The standard deviations of $\Psi$ along GL for each perturbation point for the case of $\sigma_{p1}$ (a) and $\sigma_{p2}$ (b).}
	\label{fig5}
\end{figure}

One of the possible reasons for us seeing such direction-dependent correlations might be due to the perturbation propagation features on ice shelves. The energy of perturbation propagates with the group velocity if we decompose it using Fourier transform. A similar example can be found in \citep{gudmundsson2003}. Using very simplified geometry \cite{gudmundsson2003} analyzed the propagation of basal perturbation to glacier surface and found that the direction of group velocity is aligned closely with the main flow. The existence of preferred propagation direction for the perturbations can possibly lead to our findings that favor the first principle stresses ({\bf{TZ: I am still not able to explain why it's exactly the first principle stress direction :(}}).

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig6_new.pdf}
    \caption{An example showing the relationship of the change of velocity (blue) with $\sigma_{p1}$ (red; a) and $\sigma_{p2}$ (red; b). The $x$-axis shows the cell number along GL. The inset shows the perturbation location.}
	\label{fig6}
\end{figure}

\subsection{Application on Larsen C ice shelf}

To validate if there are the same $N_b$-$N_r$ patterns as in the MISMIP+ case for realistic ice shelves, we apply the same analysis processes as above to the Larsen C ice shelf. Because the mesh resolution varies from GL to calving front, we apply 20 km $\times$ 20 km square boxes to the perturbation experiments and adjust the box areas by counting actual cell numbers that each box includes. In addition, to account for the complex geometry (GL shape and ice rises) of Larsen C ice shelf, we use two different perturbation sets with and without including near-GL boxes for investigating further the nonlinearity feature we observe in the above MISMIP+ section.

From Figure \ref{fig7}a, we can see similar $N_b$-$N_r$ correlation patterns to that of the MISMIP+ case. Along the $\sigma_{p1}$ direction we have the best $N_b$-$N_r$ correlations ($\Delta\phi=0$, red curve), whereas along the $\sigma_{p2}$ ($\Delta\phi=90$, red curve) direction they are the most insignificant. The phase difference between the $\sigma_{p1}$ and flow direction results can also be partially explained by their respective angle differences. In Figure \ref{fig7}b we can find the angles of flow directions are mostly around 90--100 degrees larger than that of the $\sigma_{p1}$ directions. This is a bit biased than the around 70 degree difference in Figure \ref{fig7}a. However, considering the stress values (and thus $N_b$) for each perturbation box are averaged over multiple cells, we argue this difference is probably an acceptable error during our calculation.

By including some near-GL perturbation boxes (Fig. \ref{fig7}c and d), the $r^2$ features get disturbed as well. Although the angle differences between the flow and $\sigma_{p1}$ directions are still mostly near 90 degree, there are no longer clear patterns in the phase differences for the corresponding $r^2$ curves. In addition, the $\sigma_{p1}$ directions are also no longer indicating the best $r^2$ correlations. However, the directions along $\sigma_{p2}$ appears to still pointing to the weakest $N_b$-$N_r$ correlations, despite the overall much smaller $r^2$ values in this case.

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig7_new.pdf}
    \caption{(a, c) The $N_b$-$N_r$ regression coefficients for each direction rotated anti-clock-wisely from the $\sigma_{p1}$ (red) and flow direction (blue); (b, d) The histogram of the angle differences between flow and $\sigma_{p1}$ directions. The insets in b) and d) show the perturbation boxes (blue circles).}
	\label{fig7}
\end{figure}

\subsection{Impacts of near-GL perturbations}
For the near-GL perturbations, it is hard to find similar linear regression relationship as discussed above. Alternatively, they are largely controlled by the distance between GL and perturbation points and also by the geometric features around them. As the perturbation decays over distance \citep{lick1970}, the neighboring GL cells of those near-GL perturbations will relatively easily detect the perturbation energy. This can be verified by looking at the standard deviations of GL velocity change due to each perturbation (Fig \ref{fig8}). For perturbations close to GL, their corresponding GL flux changes are in general confined to local regions, while in the remote GL sections the velocity changes are often negligible, resulting in large standard deviations. This can possibly cause spatial heterogeneity of GL retreating if the sub-shelf melting is very close to GL and is heavily local confined.

The propagation of perturbation can also be impacted by the spatial GL geometry, e.g., they can be blocked by the local GL. For example, the perturbation at around x = 480 km and y = 65 km in Figure \ref{fig8}a can not directly impact the ice flow on the other side of the grounded peninsula (e.g., x=485 km, y=70km) in the same way as for it's neighboring cells. This is one of the major factors that complicate our diagnostic analysis for real ice shelves (for example, Larsen C) containing complex GL shapes and geometries.

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig8.pdf}
    \caption{Standard deviation of velocity change along GL for each perturbation point for the MISMIP+ (a) and Larsen C (b) experiment. The red dashed lines (points) are the GLs for MISMIP+ (a) and Larsen C (b) .}
	\label{fig8}
\end{figure}

\subsection{Adjoint sensitivity}

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig9.pdf}
    \caption{(a, c) Perturbation points. The red points indicate near-GL ($<$2 km) points. GL is the black curve; (b, d) Model sensitivity by perturbation experiments versus adjoint sensitivity. The red dashed curve shows their linear regression.}
	\label{fig9}
\end{figure}

Need helps from Steve and Mauro!

%\subsection{MISMIP+}
%Before and after the perturbation, the GL velocity change clearly shows a spatial pattern (Fig \ref{stndVarFluxP}a). By calculating the standard deviation of the velocity change along GL for each 2 km perturbation experiment, we can see that the GL velocity change data is more dispersed for perturbation near GL, while for pertuirbations at the central region of the ice shelf, the corresponding GL velocity change is less variant along GL. If we look at Figure \ref{stndVarFluxP}b, the integrated GL flux is more contributed by the GL cells that close to the perturbation spot, i.e., the neighboring GL cells appears to have more flux contribution weight than those far away from the perperturbation location, and the perturbations near the ice shelf center tend to impact more uniformly on the GL dynamics.

%\begin{figure}
%\centering
%\includegraphics[width=1\linewidth]{figs/stndVarFluxP.pdf}
%\caption{(a) The distribution of the standard variance of the ice velocity changes on the grounding line of MISMIP+ for all perturbation experiments (note we only do the perturbation experiments for half space in $y$ of MISMIP+); (b) We find five cells on the grounding line that are the nearset to the perturbation spot. The x-axis is their mean distances to the perturbation box. The y-axis shows their combined flux contribution in the total integrated grounding line flux.}
%\label{stndVarFluxP}
%\end{figure}

%This can be further examined by looking at the relationship between GL flux response number ($N_r$) and the buttressing number ($N_b$) for the first principle stress ($\sigma_{p1}$) at different $dG$ and $dC$ (Fig \ref{diffdCdG}). $dG$ ($dC$) is the minimum distance between perturbation boxes and GL (calving front). When both $dG$ and $dC$ increase, perturbation boxes are farther away from GL and calving front, the velocity changes at GL become less dispersible and their relationship gets less ``noisy'' and tends to show linear trends. The possible explanaions can be found in the following ``Discussions'' section.

%\begin{figure}
%\centering
%\includegraphics[width=1\linewidth]{figs/diffdCdG.pdf}
%\caption{The relationships between the groundling line flux response number ($N_r$) and the buttressing number ($N_b$) at the perturbation spots by different geometric filters of $dG$ and $dC$. Larger $dG$ and $dC$ give perturbation spots that are more centralized in the ice shelf. Smaller $dG$ and $dC$ mean the perturbation locations are either close to the grounding line or the calving front. The color bars show the values of $dG$ for each perturbation spot.}

%\label{diffdCdG}
%\end{figure}

%In fact, $\sigma_{p1}$ appears to be the best metric for predicting such linear relationship, comparing to the other 7 stress components (Fig \ref{diffStressComp}). The $N_b$ for $\sigma_f$, $\sigma_{af}$, $\sigma_x$ and $\sigma_y$ show some weak correlations with GL flux response number $N_b$. The shear stress and hoop stress, however, have no clear evidences of relating to $N_r$. Note that for all plots in Figure \ref{diffStressComp}, we have removed the ``noisy'' perturbation spots using $dG$ and $dC$. It is suprising that $\sigma_{p1}$ shows better performance than $\sigma_{p2}$ and $\sigma_s$, as $\sigma_{p2}$ gives the maximum buttressing number and predicts well the passive ice regions in \cite{furst2016} and the shear stress supply the main large resistence force for divergent ice flow regions (most of the MISMIP+ ice shelf) \citep{wearing2016}.

%\begin{figure}
%\centering
%\includegraphics[width=1\linewidth]{figs/diffStressComp.pdf}
%\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the MISMIP+ 2 km $\times$ 2 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
%\label{diffStressComp}
%\end{figure}

%The buttressing number for $\sigma_{p1}$ also performs well for the 10 km $\times$ 10 km square box perturbation experiments (Fig \ref{diffStressCompAvg}), for which we average the stresses across all cells in each square box. In fact, except the shear stress $\sigma_s$, all correlations are improved compared to the 2 km $\times$ 2 km perturbation experiments. In particular, the hoop stress shows some improved predictability as well. But again, $\sigma_{p2}$ is still not preferable than $\sigma_{p1}$.

%\begin{figure}
%\centering
%\includegraphics[width=1\linewidth]{figs/diffStressCompAvg.pdf}
%\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the MISMIP+ 10 km $\times$ 10 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
%\label{diffStressCompAvg}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=1\linewidth]{figs/diffStressCompAvgL.pdf}
%\caption{The relationships between grounding line flux response number ($N_r$) and eight different stress components, $\sigma_f$ (a), $\sigma_{af}$ (b), $\sigma_{p1}$ (c), $\sigma_{p2}$ (d), $\sigma_{x}$ (e), $\sigma_{y}$ (f), $\sigma_{s}$ (g) and $\sigma_h$ (h). See Table \ref{stressDef} for the definition of each stress component in the Larsen C 20 km $\times$ 20 km box experiments. The color bars show the values of $dG$ for each perturbation spot.}
%\label{diffStressCompAvgL}
%\end{figure}

@@

%\section{Discussions}
%
%Based on the above discussion we hypothesize that the two main factors controlling how a local perturbation in ice shelf thickness leads to changes in ice flux across the GL are 1) the distance between a perturbation location and the GL and 2) the first principle stress at the perturbation location. Figure \ref{distPerturbComp} shows an example of the impact of the perturbation location on the velocity changes across the ice shelf and GL. Despite the complex patterns of velocity change around the perturbation spot, we can see two clear features: 1) the magnitude of velocity change decreases with distances away from the perturbation spot and 2) the velocity change pattern becomes different and more disordering when the perturbation  approaches to GL. 
%
%
%The first feature is easy to understand because the energy of perturbation decays as it propagate away to far upstream and downstream. However, the perturbation propagation is not simply dependent on distance, but replies on the ice flow field. A complete perturbation propagation theory is necessary to describe/explain these small velocity changes across the ice shelf, however, it is beyond the scope of this study. The second one can probably be explained by two ice-flow mechanisms: 1) ice shelf dynamics are remarkably impacted by the grounded ice flow near the GL since it's a coupled system for grounded ice sheet and floating ice shelf and in MALI they is resolved as an integrated system as well; 2) the propagation of perturbation can be impacted by the spatial GL geometry. For example, the perturbation in Figure \ref{distPerturbComp}a can not directly impact the ice on the other side of the grounded peninsula in the same way as for it's neighboring cells.
%
%In addition, the impacts of distance can also be verified from the box averaging experiments (Figure \ref{diffStressCompAvg} and \ref{diffStressCompAvgL}). By smoothing out the stresses in the perturbation box, different distances to GL for each cells and their impacts in the perturbation box are also averaged, resulting in better linear $N_b$-$N_r$ relationships. This possibly indicates that those linear $N_b$-$N_r$ relationships can apply for realistic ice shelves, as in reality the basal melting always occurs in wide spatial regions.
%
